
Installing the libraries
```{r}
library(tidyverse)
library(spatstat)
```

Downloading the spatial centered data of specific cell type
```{r}
coordinates <- read.csv("cell_type_coordinates.csv",
                        row.names = 1) # data for cancer cells
```

```{r}
dim(coordinates)
```

```{r}
summary(coordinates)
```

Convertion the coordinates from the data to matrix
```{r}
cell_mat <- as.matrix(coordinates[1:3])
```

### Let the spatial statistics begin!
Define the ranges for each coordinate (you could also use range(cell_mat[,i]))
```{r}
pp_box <- function(mat) {
  x_range <- c(min(mat[, 1]), max(mat[, 1]))
  y_range <- c(min(mat[, 2]), max(mat[, 2]))
  z_range <- c(min(mat[, 3]), max(mat[, 3]))

  # Create the 3D box (an object of class "box3")
  box3(xrange = x_range, yrange = y_range, zrange = z_range)
}
```

Making a point pattern object
```{r}
mypattern <- pp3(cell_mat[, 1], cell_mat[, 2],
                 cell_mat[, 3], pp_box(cell_mat))
```

```{r}
plot(mypattern)
```

Ripley-K function
```{r}
# Compute the 3D Ripley's K function for your pp3 object
k3 <- K3est(mypattern)

# Plot the estimated K function
plot(k3)
```

Pair-correlation function
```{r}
# Compute the 3D pair correlation function for your pp3 object
pcf3 <- pcf3est(mypattern)

# Plot the estimated pair correlation function
plot(pcf3)
```


## Compute dissections of the 3D point pattern
### Declare the rotation matrices
```{r}
rotation <- function(a, b, g) {
  matrix(c(
           cos(a) * cos(b),
           cos(a) * sin(b) * sin(g) - sin(a) * cos(g),
           cos(a) * sin(b) * cos(g) + sin(a) * sin(g),
           sin(a) * cos(b),
           sin(a) * sin(b) * sin(g) + cos(a) * cos(g),
           sin(a) * sin(b) * cos(g) - cos(a) * sin(g),
           -sin(b),
           cos(b) * sin(g),
           cos(b) * cos(g))
  , nrow = 3, byrow = TRUE)
}
```

### Declare the slice function
```{r}
slice <- function(a, b, c, slice_size, mut_frame) {
  "Compute the slice of the 3D point pattern."
  value <- a * mut_frame[, 1] + b * mut_frame[, 2] + c * mut_frame[, 3]
  # Return a logical vector: TRUE if value is within the slice, FALSE otherwise.
  value >= -slice_size / 2 & value <= slice_size / 2
}
```

```{r}
coordinates %>%
  filter(slice(0, 0, 1, 2, as.matrix(coordinates[, 1:3])))
```

### Define the function to compute the dissections
```{r}
pc_for_slice <- function(g, b, a, cell_mat) {
  "Compute the dissections of the 3D point pattern."
  mut_mat <- cell_mat %*% rotation(g, b, a)
  # Convert mutation_matrix to dataframe object to apply slice function
  mut_frame <- as.data.frame(mut_mat)
  # Create a slice of the 3D point pattern
  slices <- mut_frame |>
    (\(df) filter(df, slice(0, 0, 1, 2, as.matrix(select(df, X, Y, Z)))))()
  # Create a 3D point pattern object
  pp <- pp3(slices, y, z, box)
  # Compute the 3D pair correlation function
  pcf3est(pp)
}
```

```{r}
pp_box(cell_mat)
```